Estimating Predictive Rate-Distortion Curves via Neural Variational Inference

### 2019-04-23 paper.tex

\date{2018}

@@

% Authors, for the paper (add full first names)
\Author{Michael Hahn $^{1}$, and Richard Futrell $^{2}$}

@@

\rljf{Small formatting changes I made: \\
* I prefer bold over italics when introducing terms, because when someone is visually scanning the paper looking for a definition, bold stands out much better. \\
* Punctuation should be attached to equations as if they were part of the text (so sometimes periods and commas after equations) \\
* no spaces between m-dashes---like this \\
* we should make sure not to use ``this'', ``that'', ``which'', etc. as bare NPs. Copyeditors don't like this.
}

@@

\paragraph{NPRD}
For all experiments, we used $M=15$.
Neural networks have hyperparameters, such as the number of units and the step size used in optimization, which affect the quality of approximation depending on properties of the dataset and the function being approximated.
Given that NPRD provably provides upper bounds on the PRD objective~(\ref{eq:prd-loss}), one can in principle identify the best hyperparameters for a given process by choosing the combination that leads to the lowest estimated upper bounds.
As a computationally more efficient method, we defined a range of plausible hyperparameters based both on experience reported in the literature, and considerations of computational efficiency.
These parameters are shown in Table~\ref{tab:nprd-hyperparameters}.
We then randomly sampled, for each of the processes that we experimented on, combinations of $\lambda$ and these hyperparameters to run NPRD on.

@@

As $M \rightarrow \infty$, (\ref{eq:sol}) behaves as
\begin{equation}
	P \approx E_0 - \alpha\beta \cdot (R/\alpha)^{\frac{\beta-1}{\beta}} e^{-(R/\alpha)^{1/\beta}},
\end{equation}
which asymptotes to $E_0$.
Given the values fitted above ($\alpha=1$, $\beta=1.7$), we found that $E_0 = 1.13$ yielded a good fit.
Using~(\ref{eq:r-alpha-beta}), this can be extended without further parameters to the third curve in Figure~\ref{fig:eng-pos}.
Resulting fits are shown in Figure~\ref{fig:eng-pos-fitted}.
%$E = E_0 - \Gamma(1.6, (\log\beta))$

@@

Our analysis of PRD curves for natural language suggests that it is characterized by very high statistical complexity beyond its excess entropy.

\citet{dkebowski2018natural} argues that the excess entropy of connected texts in natural language is unbounded, meaning that the statistical complexity is unbounded. Our result suggests that the statistical complexity within individual sentences is unbounded (though the excess entropy is finite). Together, these results

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided. The following statements should be used “conceptualization, X.X. and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y. and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing—original draft preparation, X.X.; writing—review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y.”, please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation. Authorship must be limited to those who have contributed substantially to the work reported.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
\appendixsections{multiple} %Leave argument "multiple" if there are multiple sections. Then a counter is printed ("Appendix A"). If there is only one appendix section then change the argument to "one" and no counter is printed ("Appendix").
\appendix
\section{Hyperparameters}

@@

\begin{table}
\begin{tabular}{l||lll}
	& Section~\ref{sec:tractable} &   Section~\ref{sec:pos} & Section~\ref{sec:words}\\ \hline\hline
	Embedding Dimension & 50, 100 &    50, 100, 200, 300   &  150      \\
	LSTM Dimension & 32           &   64, 128, 256, 512    &    512     \\ \hline
%	LSTM Layers & 1 \\ \hline
	Dropout rate & 0.0, 0.1, 0.2  &   0.0, 0.1, 0.2    &      0.4     \\
	Input Dropout & 0  &    0.0, 0.1, 0.2   &        0.2             \\ \hline
	Adam Learning Rate & $\{1,5\} \cdot 10^{-4}$, $\{1,2,4\} \cdot 10^{-3}$  &   $\{1,5\} \cdot 10^{-4}$, $\{1,2,4\} \cdot 10^{-3}$ &       0.00005, 0.0001, 0.0005, 0.001     \\
	Batch Size & 16, 32, 64  &    16, 32 64   &       32          \\ \hline
	Flow Length & 1, 2  &   1,2,3,4,5    &          4              \\
	Flow Type & DSF, DDSF  &    DSF, DDSF   &       DSF, DDSF             \\
	Flow Dimension & 32, 64, 128, 512  &   512    &    512        \\
	Flow Layers & 2  &    2   &                 2         \\
\end{tabular}
	\caption{NPRD Hyperparameters}\label{tab:nprd-hyperparameters}
\end{table}

\unskip
\subsection{}
The appendix is an optional section that can contain details and data supplemental to the main text. For example, explanations of experimental details that would disrupt the flow of the main text, but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data is shown in the main text can be added here if brief, or as Supplementary data. Mathematical proofs of results not central to the paper can be added as an appendix.

\section{}
All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\sampleavailability{Samples of the compounds ...... are available from the authors.}

>>>>

@@

% Authors, for the paper (add full first names)
\Author{Michael Hahn $^{1}$ and Richard Futrell $^{2}$}

@@

@@

\paragraph{NPRD}
For all experiments, we used $M=15$.
Neural networks have hyperparameters, such as the number of units and the step size used in optimization, which affect the quality of approximation depending on properties of the dataset and the function being approximated.
Given that NPRD provably provides upper bounds on the PRD objective~(\ref{eq:prd-loss}), one can in principle identify the best hyperparameters for a given process by choosing the combination that leads to the lowest estimated upper bounds.
As a computationally more efficient method, we defined a range of plausible hyperparameters based both on experience reported in the literature, and considerations of computational efficiency.
These parameters are discussed in Appendix~\ref{sec:app-hyperparams}.
We then randomly sampled, for each of the processes that we experimented on, combinations of $\lambda$ and these hyperparameters to run NPRD on.

@@

As the window size $M \rightarrow \infty$, (\ref{eq:sol}) behaves as
\begin{equation}
	P \approx E_0 - \alpha\beta \cdot (R/\alpha)^{\frac{\beta-1}{\beta}} e^{-(R/\alpha)^{1/\beta}},
\end{equation}
which asymptotes to $E_0$.
Given the values fitted above ($\alpha=1$, $\beta=1.7$), we found that $E_0 = 1.13$ yielded a good fit.
Using~(\ref{eq:r-alpha-beta}), this can be extended without further parameters to the third curve in Figure~\ref{fig:eng-pos}.
Resulting fits are shown in Figure~\ref{fig:eng-pos-fitted}.
%$E = E_0 - \Gamma(1.6, (\log\beta))$

@@

Our analysis of PRD curves for natural language suggests that it is characterized by very high and perhaps infinite statistical complexity beyond its excess entropy. In a similar vein, \citet{dkebowski2018natural} has argued that the excess entropy of connected texts in natural language is infinite (in contrast, our result is for isolated sentences). If the statistical complexity of natural language is indeed infinite, then statistical complexity is not a good complexity metric for natural language.

We suggest that the PRD curve may form a more natural complexity metric for highly complex processes such as language. Among those processes with infinite statistical complexity, some will have a gentle PRD curve---meaning that they can be well-approximated at low rates---while others will have a steep curve, meaning they cannot be well-approximated at low rates. We conjecture that although natural language may have infinite statistical complexity, it has a gentler PRD curve than other processes with this property, meaning that achieving a reasonable approximation of the predictive distribution does not require inordinate memory resources.

@@

All code for reproducing the results in this work is available at \url{https://github.com/m-hahn/predictive-rate-distortion}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{M.H. and R.F. conceived of and designed the experiments. M.H. performed the experiments and analysed data. M.H. and R.F. wrote the paper. All authors have read and approved the final manuscript.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
\appendixsections{multiple} %Leave argument "multiple" if there are multiple sections. Then a counter is printed ("Appendix A"). If there is only one appendix section then change the argument to "one" and no counter is printed ("Appendix").
\appendix
\section{Hyperparameters}\label{sec:app-hyperparams}

@@

\begin{table}
\begin{tabular}{l||lll}
	& Section~\ref{sec:tractable} &   Section~\ref{sec:pos} & Section~\ref{sec:words}\\ \hline\hline
	Embedding Dimension & 50, 100 &    50, 100, 200, 300   &  150      \\
	LSTM Dimension & 32           &   64, 128, 256, 512    &   256, 512     \\ \hline
%	LSTM Layers & 1 \\ \hline
	Dropout rate & 0.0, 0.1, 0.2  &   0.0, 0.1, 0.2    &   0.1,   0.4     \\
	Input Dropout & 0  &    0.0, 0.1, 0.2   &        0.2             \\ \hline
	Adam Learning Rate & $\{1,5\} \cdot 10^{-4}$, $\{1,2,4\} \cdot 10^{-3}$  &   $\{1,5\} \cdot 10^{-4}$, $\{1,2,4\} \cdot 10^{-3}$ &       0.00005, 0.0001, 0.0005, 0.001     \\
	Batch Size & 16, 32, 64  &    16, 32 64   & 16, 32, 64  \\ \hline
	Flow Length & 1, 2  &   1,2,3,4,5    &          1,2,3,4              \\
	Flow Type & DSF, DDSF  &    DSF, DDSF   &       DSF, DDSF             \\
	Flow Dimension & 32, 64, 128, 512  &   512    &    512        \\
	Flow Layers & 2  &    2   &                 2         \\
\end{tabular}
	\caption{NPRD Hyperparameters. See Appendix~\ref{sec:app-hyperparams} for description of the parameters.}\label{tab:nprd-hyperparameters}
\end{table}

%\unskip
%\subsection{}
%The appendix is an optional section that can contain details and data supplemental to the main text. For example, explanations of experimental details that would disrupt the flow of the main text, but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data is shown in the main text can be added here if brief, or as Supplementary data. Mathematical proofs of results not central to the paper can be added as an appendix.

%\section{}
%All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\sampleavailability{Code is available.}
